---
title: "bihar TV penetration"
output:
  html_document: default
  pdf_document: default
date: '2022-07-21'
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir="C:\\Users\\AG\\Desktop\\TV penetration in Bihar and Gujarat\\Bihar")

```

```{r,echo=F,eval=TRUE,warning=FALSE,message=FALSE}
library(car)
library(caret)
library(nortest)
library(flextable)
library(officer)
library(dplyr)
library(corrgram)
library(data.table)
library(ggplot2)
```

### Background: The data has 44098 cases for state Bihar where each case is a area in the state. The “percentage of households having TV” is a target variable. The data has many other variables related to condition of houses, ownership status,facilities etc.which can influence the TV penetration in the specific area.
### The objective is to identify significant factors affecting TV penetration.


&nbsp;
&nbsp;
&nbsp;


#### Data Description
```{r,warning=FALSE,message=FALSE,echo=F,eval=TRUE,ft.align="left",comment=""}

datadesc<-data.frame(liv_house="Percentage of total houses having livable condition",
                     owned="Percentage of owned houses ",
                     elect_light="Percentage of houses having main source of light as electricity",
                     lat_inprem="Percentage of houses having latrine facility",
                     perm="Percentag of permanent houses",
                     atlone_rm="Percentage of houses having atleast one dwelling room",
                     atlthr_mem="Percentage of houses having atleast three members",
                     atlone_cou="Percentage of houses having atleast one couple",
                     bath="Percentage of houses having bathroom with or without roof",
                     kitchen="Percentage of houses having kitchen inside or outside house")
datadesc<-as.data.frame.matrix(t(datadesc))
datadesc$Col<-row.names(datadesc)
datadesc<-datadesc[,2:1]
names(datadesc)<-c("Column name","Column description")


std_border = fp_border(color="black", width = 1)
datadesc %>% flextable() %>% 
  border_inner(border = std_border) %>%
  border_outer(border = std_border) %>%
  bold(bold = T,part = "header") %>%
  autofit()
```


#### Import data of Bihar state combining all 38 districts

```{r,warning=FALSE,message=FALSE,echo=TRUE,eval=TRUE,ft.align="left",comment=""}
library(readxl)
data=read_excel("Araria filtered.xlsx")

data.files = list.files(pattern = "*.xlsx")

i=2
lst<-list()
for (i in 2:length(data.files)) {
  lst[[i]]<- read_excel(data.files[i],skip=6)
  names(lst[[i]])<-names(data)
}                                 

lst2<-lst[lapply(lst,length)>0]

bihardata<-do.call(rbind,lst2)
bihardata<-rbind(data,bihardata)
head(data.frame(bihardata))
```


#### Creating columns atlone_rm by adding columns one room and above, atlthr_mem by adding columns three members and above , atlone_cou by adding columns one couple and above , bath by adding columns bathroom with roof and without roof, kitchen by adding columns cooking inside and outside house 

```{r,warning=FALSE,message=FALSE,echo=TRUE,eval=TRUE,ft.align="left",comment=""}
bihardata$atlone_rm=rowSums(bihardata[,c("one_rm","two_ rm","thr_rm","four_rm","five_rm","msix_rm")])


bihardata$atlthr_mem=rowSums(bihardata[,c("thr_mem","four_mem","five_mem","sixeig_mem","mnine_mem")])


bihardata$atlone_cou=rowSums(bihardata[,c("one_cou","two_cou","thr_cou","four_cou","mfive_cou")])


bihardata$bath=rowSums(bihardata[,c("y_bath","yn_bath")])

bihardata$kitchen=rowSums(bihardata[,c("yin_kit","yout_kit")])

```

#### Identify the variables and produce correlation matrix
```{r,warning=FALSE,message=FALSE,echo=TRUE,eval=TRUE,ft.align="left",comment=""}
library(reshape2)
cormat <- round(cor(bihardata[,c("liv_house","owned","elect_light","lat_inprem","perm","atlone_rm","atlthr_mem","atlone_cou","bath","kitchen","tv")]),2)
melted_cormat <- melt(cormat)

ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile()+
  geom_text(aes(Var2, Var1, label = value), color = "black", size = 4)+
  scale_fill_gradient2(low="red",mid="white",high="blue")

```

#### Split original data into training and testing data sets for Bihar state
```{r,warning=FALSE,message=FALSE,echo=TRUE,eval=TRUE,ft.align="left",comment=""}
index<-createDataPartition(bihardata$tv,p=0.8,list=FALSE) 

traindata<-bihardata[index,]
testdata<-bihardata[-index,]   

```

#### Linear Regression using lm function

```{r,warning=FALSE,message=FALSE,echo=TRUE,eval=TRUE,ft.align="left",comment=""}
hp_model<-lm(tv~liv_house+owned+elect_light+lat_inprem+perm+atlone_rm+atlthr_mem+atlone_cou+bath+kitchen,data=traindata)  
```

#### Display parameter estimates with other model statistics
```{r,warning=FALSE,message=FALSE,echo=TRUE,eval=TRUE,ft.align="left",comment=""}
summary(hp_model)
```

#### Comment:"atlone_rm" and "atlone_cou" are insignificant variables
#### The model explains 70% of variation in dependent variable

&nbsp;
&nbsp;

#### Check for multicollinearity using vif function
```{r,warning=FALSE,message=FALSE,echo=TRUE,eval=TRUE,ft.align="left",comment=""}
vif(hp_model)
```

#### Comment : As all VIF's are less than 5 ,Multicollinearity is not present in a model

&nbsp;

#### Plot of Residuals vs Predicted values

```{r,warning=FALSE,message=FALSE,echo=TRUE,eval=TRUE,ft.align="left",comment=""}
traindata$fit<-fitted(hp_model)
traindata$resi<-residuals(hp_model)
plot(traindata$fit,traindata$resi) 
```

#### Comment: It is observed that residuals are randomly distributed and uncorelated with predicted values

&nbsp;
&nbsp;

#### Check if distribution of errors is "NORMAL"
```{r,warning=FALSE,message=FALSE,echo=TRUE,eval=TRUE,ft.align="left",comment=""}
qqnorm(traindata$resi) 
```

```{r,warning=FALSE,message=FALSE,echo=TRUE,eval=TRUE,ft.align="left",comment=""}
lillie.test(traindata$resi)
```

#### Comment: Although normality of errors is not established we will proceed to evaluate the model performance
&nbsp;

#### Rerun the model after removing the insignificant variables

```{r,warning=FALSE,message=FALSE,echo=TRUE,eval=TRUE,ft.align="left",comment=""}
hp_model2<-lm(tv~liv_house+owned+elect_light+lat_inprem+perm+atlthr_mem+bath+kitchen,data=traindata) 
summary(hp_model2)
```


#### Calculate RMSE values based on residuals using first principle
```{r,warning=FALSE,message=FALSE,echo=TRUE,eval=TRUE,ft.align="left",comment=""}
traindata$resi<-residuals(hp_model2)

RMSE<-sqrt(mean(traindata$resi**2))
RMSE
```
#### Model Validation: Holdout Method using RMSE

```{r,warning=FALSE,message=FALSE,echo=TRUE,eval=TRUE,ft.align="left",comment=""}
testdata$pred<-predict(hp_model2,testdata)
testdata$res<-(testdata$tv-testdata$pred)
RMSEtest<-sqrt(mean(testdata$res**2))
RMSEtest
```

#### Comment: RMSE values for train and test data are similar hence the model can be implemented for decision making for Bihar state

&nbsp;
&nbsp;

#### Converting continuous variables into categorical variables by using quantile function

```{r,warning=FALSE,message=FALSE,echo=T,eval=TRUE,ft.align="left",comment=""}
bihardata$TV<-cut(bihardata$tv,quantile(bihardata$tv),include.lowest=TRUE,labels=c("very low","low","moderate","high"))
bihardata$livable<-cut(bihardata$liv_house,quantile(bihardata$liv_house),include.lowest=TRUE,labels=c("very low","low","moderate","high"))

bihardata$owner <- as.factor(ntile(bihardata$owned, 4))
levels(bihardata$owner) <- c("very low","low","moderate","high")

bihardata$electr <- as.factor(ntile(bihardata$elect_light, 4))
levels(bihardata$electr) <- c("very low","low","moderate","high")

bihardata$latrine<-cut(bihardata$lat_inprem,quantile(bihardata$lat_inprem),include.lowest=TRUE,labels=c("very low","low","moderate","high"))
bihardata$perman<-cut(bihardata$perm,quantile(bihardata$perm),include.lowest=TRUE,labels=c("very low","low","moderate","high"))
bihardata$membr<-cut(bihardata$atlthr_mem,quantile(bihardata$atlthr_mem),include.lowest=TRUE,labels=c("very low","low","moderate","high"))
bihardata$bathrm<-cut(bihardata$bath,quantile(bihardata$bath),include.lowest=TRUE,labels=c("very low","low","moderate","high"))
bihardata$kit<-cut(bihardata$kitchen,quantile(bihardata$kitchen),include.lowest=TRUE,labels=c("very low","low","moderate","high"))

bihar1=data.frame(bihardata$TV,bihardata$livable,bihardata$owner,bihardata$electr,bihardata$latrine,bihardata$perman,bihardata$membr,bihardata$bathrm,bihardata$kit)
names(bihar1)=c("TV","livable","owner","electr","latrine","perman","membr","bathrm","kit")

lst<-list()
i=1
for (i in 1:8){
 
  a<-as.data.frame.matrix(table(bihar1$TV,bihar1[,i+1]))
  a2<-as.data.frame.matrix(round(prop.table(table(bihar1$TV,bihar1[,i+1]))*100,2))
  for (j in 1:4){
    a[,j]<-paste(a[,j]," (",a2[,j],"%)",sep = "")
  }
  a$New<-c("very low","low","moderate","high")
  
  a<-a[,c(5,1:4)]
  names(a)[1]<-paste("TV & ",names(bihar1)[i+1],sep = "")
  lst[[i]]<-a
  names(lst)[i]<-names(bihar1)[i+1]
  
}

std_border = fp_border(color="black", width = 1)
lst[[1]] %>% flextable() %>% 
  set_caption(caption = "TV vs liv_house",autonum = NULL,
  style = "Table Caption",
  html_escape = TRUE) %>%
  border_inner(border = std_border) %>%
  border_outer(border = std_border) %>%
  bold(bold = T,part = "header") %>%
  autofit()

lst[[2]] %>% flextable() %>% 
  set_caption(caption = "TV vs Owned",autonum = NULL,
  style = "Table Caption",
  html_escape = TRUE) %>% 
  border_inner(border = std_border) %>%
  border_outer(border = std_border) %>%
  bold(bold = T,part = "header") %>%
  autofit()

lst[[3]] %>% flextable() %>% 
  set_caption(caption = "TV vs elect_light",autonum = NULL,
  style = "Table Caption",
  html_escape = TRUE) %>% 
  border_inner(border = std_border) %>%
  border_outer(border = std_border) %>%
  bold(bold = T,part = "header") %>%
  autofit()

lst[[4]] %>% flextable() %>% 
  set_caption(caption = "TV vs lat_inprem",autonum = NULL,
  style = "Table Caption",
  html_escape = TRUE)  %>%
  border_inner(border = std_border) %>%
  border_outer(border = std_border) %>%
  bold(bold = T,part = "header") %>%
  autofit()

lst[[5]] %>% flextable() %>% 
  set_caption(caption = "TV vs perm",autonum = NULL,
  style = "Table Caption",
  html_escape = TRUE) %>%
  border_inner(border = std_border) %>%
  border_outer(border = std_border) %>%
  bold(bold = T,part = "header") %>%
  autofit()

lst[[6]] %>% flextable() %>% 
  set_caption(caption = "TV vs atlthr_mem",autonum = NULL,
  style = "Table Caption",
  html_escape = TRUE) %>%
  border_inner(border = std_border) %>%
  border_outer(border = std_border) %>%
  bold(bold = T,part = "header") %>%
  autofit()

lst[[7]] %>% flextable() %>% 
  set_caption(caption = "TV vs bath",autonum = NULL,
  style = "Table Caption",
  html_escape = TRUE) %>%
  border_inner(border = std_border) %>%
  border_outer(border = std_border) %>%
  bold(bold = T,part = "header") %>%
  autofit()

lst[[8]] %>% flextable() %>% 
  set_caption(caption = "TV vs kitchen",autonum = NULL,
  style = "Table Caption",
  html_escape = TRUE) %>%
  border_inner(border = std_border) %>%
  border_outer(border = std_border) %>%
  bold(bold = T,part = "header") %>%
  autofit()

```

#### Plot quantiles of independent variables against high values of TV penetration
```{r,warning=FALSE,message=FALSE,echo=T,eval=TRUE,ft.align="left",comment=""}
lst2<-list()
i=1
for (i in 1:8){
  a<-as.data.frame.matrix(table(bihar1$TV,bihar1[,i+1]))
  a$New<-row.names(a)
  try<-subset(a,a$New=="high",select = c("very low","low","moderate","high"))
  try$new<-names(lst)[i]
  lst2[[i]]<-try
}

trydata<-do.call(rbind,lst2)

library(reshape2)
trydata<-melt(trydata)
names(trydata)<-c("Variable","Quantile","Value")

library(plyr)
trydata = ddply(trydata, .(Variable), transform, percent = Value/sum(Value) * 100)
trydata$label = paste(round(trydata$percent,2), "%",sep = "")

library(ggplot2)
ggplot(trydata, aes(fill=Quantile, y=Value, x=Variable,label=Value)) + 
  geom_bar(position="stack", stat="identity")+
  geom_text(aes(label=label),size = 3, position = position_stack(vjust = 0.5))


```

